Tidy Data

### 2014-01-10 tidy-data.tex
2.0 and 2.1

Like families, tidy datasets are all alike, but every messy dataset is messy in its own way. Tidy datasets provide a standard way to connect the structure of a dataset (its physical layout) to its semantics (its meaning). I'll provide some standard vocabulary for describing the structure of a dataset, its semantics, and then use those definitions to define tidy data.

@@

Most statistical datasets are rectangular tables made up of \textbf{columns} and \textbf{rows}. The columns are almost always labelled, and rows are often labelled. Table~\ref{tbl:preg-raw-1} provides some data about an imaginary experiment in a format commonly seen in the wild. The table has two columns and three rows, and both rows and columns are labelled.

@@

There are many ways to structure the same underlying data. Table~\ref{tbl:preg-raw-2} shows the same data as Table~\ref{tbl:preg-raw-1}, but the rows and columns have been transposed. The data is the same, but the layout is different. Our vocabulary of rows and columns is not rich enough to say why these two tables represent the same data. We also need a way to describe the meaning of the underlying data displayed in the table. It's not enough to describe the appearance of a dataset (how the values are physically laid out), we also need a way to describe the underlying semantics, the meaning, of the values in the dataset.

>>>>

Like families, tidy datasets are all alike but each is messy in its own way. Tidy datasets provide a standardized way to link the structure of a dataset (its physical layout) with its semantics (its meaning). In this section, I'll provide some standard vocabulary for describing the structure of a dataset, its semantics, and then use those definitions to define tidy data.

@@

Most statistical datasets are rectangular tables made up of \textbf{rows} and \textbf{columns}. While columns are almost always labelled, rows are sometimes not. Table~\ref{tbl:preg-raw-1} provides some data about an imaginary experiment in a format commonly seen in the wild. The table has two columns and three rows, and both rows and columns are labelled.

@@

There are many ways to structure the same underlying data. Table~\ref{tbl:preg-raw-2} shows the same data as Table~\ref{tbl:preg-raw-1}, but the rows and columns have been transposed. The data is the same, but the layout is different. Our vocabulary of rows and columns is simply not rich enough to tell us why the two tables represent the same data. In addition to appearance, what we need is a way to describe the underlying semantics, or meaning, of the values displayed in the dataset.
